---
title: "Analyzing, visualizing, modeling Kaggle Titanic Dataset in R"
output: html_document
---


 **VARIABLE DESCRIPTIONS: **


survival        :Survival

                (0 = No; 1 = Yes)
                
pclass          :Passenger Class

                (1 = 1st; 2 = 2nd; 3 = 3rd)
                
name            :Name

sex             :Sex

age             :Age

sibsp           :Number of Siblings/Spouses Aboard

parch           :Number of Parents/Children Aboard

ticket          :Ticket Number

fare            :Passenger Fare

cabin           :Cabin

embarked        :Port of Embarkation

                (C = Cherbourg; Q = Queenstown; S = Southampton)
                

 **SPECIAL NOTES: **
Pclass is a proxy for socio-economic status (SES)

 1st ~ Upper; 2nd ~ Middle; 3rd ~ Lower

Age is in Years; Fractional if Age less than One (1)

 If the Age is Estimated, it is in the form xx.5

With respect to the family relation variables (i.e. sibsp and parch)
some relations were ignored.  The following are the definitions used
for sibsp and parch.

Sibling:  Brother, Sister, Stepbrother, or Stepsister of Passenger Aboard Titanic

Spouse:   Husband or Wife of Passenger Aboard Titanic (Mistresses and Fiances
Ignored)

Parent:   Mother or Father of Passenger Aboard Titanic

Child:    Son, Daughter, Stepson, or Stepdaughter of Passenger Aboard Titanic


Other family relatives excluded from this study include cousins,
nephews/nieces, aunts/uncles, and in-laws.  Some children travelled
only with a nanny, therefore parch=0 for them.  As well, some
travelled with very close friends or neighbors in a village, however,
the definitions do not support such relations.

 **Getting Data  **

```{r,message=FALSE,warning=FALSE}
library(party)
library(randomForest)
library(rattle)
library(rpart)
library(caret)
library(dplyr)
library(lattice)
library(Hmisc)
setwd("C:/Onur_Data/Folder/DATA SCIENCE/titanic")

test<-read.csv("test.csv",header = TRUE,na.strings = c("NA",""))
train<-read.csv("train.csv",header = TRUE,na.strings = c("NA",""))
genderclassmodel<-read.csv("genderclassmodel.csv",header = TRUE,na.strings = c("NA",""))
gendermodel<-read.csv("gendermodel.csv",header = TRUE,na.strings = c("NA",""))
alldata<- bind_rows(train,test)

```


 **Finding NA Amounts with Loop Functions **


```{r}
c<-data.frame(V1=character(), 
                     V2=character(),
                     V3=character(),
                     stringsAsFactors=FALSE)
for(i in 1:length(names(train))){
    c[i,2]<-sum(is.na(train[,i]))}
for(i in 1:length(names(alldata))){
    c[i,3]<-sum(is.na(alldata[,i]))}

c$V1<-names(train)
names(c)<-c("feature","Train NA-Counts","ALL NA-Counts")
c



```

 **Dealing with NAs **



```{r}

meanage<-mean(alldata$Age,na.rm = TRUE)
meanage

train[is.na(train$Age),6]<-meanage
train$farecut<-cut2(train$Fare,c(125,250,375))
test$farecut<-cut2(test$Fare,c(125,250,375))
which(is.na(train$Embarked))
train[c(62,830),12]<-"S"
```


 **Visualizing **  
 
```{r}
histogram( ~Survived,train,breaks=seq(from=-.5,to=1.5,by=1),xlim = c(-.5,1.5))
histogram( ~Survived|factor(Sex),train,breaks=seq(from=-.5,to=1.5,by=1),xlim = c(-.5,1.5))
histogram( ~Pclass|factor(Sex)+factor(Survived),train)
histogram( ~Age|factor(Sex)+factor(Survived),train)
histogram( ~SibSp|factor(Sex)+factor(Survived),train)
histogram( ~Parch|factor(Sex)+factor(Survived),train)
histogram( ~Fare|factor(Sex)+factor(Survived),train)
histogram( ~Embarked|factor(Sex)+factor(Survived),train)
histogram( ~Embarked|factor(farecut)+factor(Survived),train)
histogram( ~Embarked|factor(Pclass)+factor(Survived),train)
```


**Modelling**

*Principal component analysis*
```{r}
pairs(train[,c(3,6,7,8,10)])
pca<-princomp(train[,c(3,6,7,8,10)],cor = TRUE,scores = TRUE)
plot(pca)
biplot(pca)
pca$loadings
head(pca$scores)
pca <- preProcess(train[,c(3,6,7,8,10)], 
                   method=c( "center", 
                     "scale", "pca"))
pc<- predict(pca, train[,c(3,6,7,8,10)])
print(pca)
summary(pca)
plot(pc)

```

*Decision Tree*
```{r}
fit<-rpart(Survived~Pclass+Sex+Age+SibSp+Parch+farecut+Embarked,data = train)
fancyRpartPlot(fit)
pred<-round(predict(fit,test))
confusionMatrix(pred,genderclassmodel$Survived)
```

*Random Forest*
```{r}
set.seed(313)
fitrf<-randomForest(as.factor(Survived)~Pclass+Sex+Age+SibSp+Parch+Fare,data = train,ntree=2000,importance=TRUE)
print(fitrf) 
print(importance(fitrf,type = 2)) 
pred2<-predict(fitrf,test)
confusionMatrix(pred2,genderclassmodel$Survived)
varImpPlot(fitrf)
```
